---
title: "On # of days of bad mental health in the past 30 days"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(ggplot2)
library(survey)
library(srvyr)
library(readr)
library(forcats)
library(gtsummary)

# Set options for allowing a single observation per stratum 
options(survey.lonely.psu = "adjust")
```

```{r include = FALSE}
# Load data
load(file = "./data/brfss_design.Rdata")
```

### Summary statistics

Below is the unweighted counts of the all included demographic variables, exposure, and other covariates, grouped by the self-reported number of days with bad mental health in the past 30 days in the following two categories: fewer than 15 days, and more than 15 days. 

```{r}
tb1_unweighted =
  brfss_design$variables %>%
  tbl_summary(by = mh_bin, percent = "row", 
              include = c(sex, race, age, marital_status,
                             children,
                             education, income, employment,
                             general_health, exercise),
              type = list(exercise ~ "categorical"))

tb1_unweighted
```

We make the following observations about the above summary counts:

* Female are more likely to report longer period with bad mental health.
* Asian are less likely to report longer period with bad mental health, while the proportion for other races are similar.
* It seems that people of older age, married, higher education level, higher income, better general health status and doing more exercise are less likely to be in bad mood for over half a month.
* It seems that the number of children does not influence the length of time people are in bad mood.
* The employed  are less likely to report longer period with bad mental health, while those are unemployed are much more likely to be in bad mood for over half a month.


```{r weighted summary}
tb1_weighted =
  brfss_design %>%
  tbl_svysummary(by = mh_bin, percent = "row", 
                 include = c(sex, race, age, marital_status,
                             children,
                             education, income, employment,
                             general_health, exercise),
                  type = list(exercise ~ "categorical"))

tb1_weighted
```

### Gender, #of days of bad mental health in the past 30 days

Before analyzing the influence of other factors on the # of days of bad mental health in the past 30 days, we would like to determine if the outcome is independent of gender or not. We do a chi-square test for this and the p-value is shown below:

```{r}
brfss_design %>%
  svychisq(~sex + mh_bin, . , statistic="adjWald") %>% 
  broom::tidy() %>% 
  pull(p.value)
```

The test statistic proved that # of days of bad mental health in the past 30 days is dependent on gender. Thus we examine the influences of other factors separately based on different genders in the following analysis.

### Gender, number of children in family, and # of days of bad mental health in the past 30 days

We are interested in the differential impact of the number of children in the household on self-reported prevalence of over 15 days of bad mental health in the past 30 days among men and women. Below is a barplot on this topic:

```{r}
brfss_design %>%
  group_by(sex, children, mh_bin) %>%
  summarize(mh_prev = 100*survey_mean(na.rm = T, vartype = c("ci"))) %>%
  filter(!is.na(children), mh_bin == ">15 days") %>%
  select(-mh_bin) %>%
  ggplot(aes(x = children, y = mh_prev, fill = sex)) +
  geom_bar(position="dodge", stat="identity") +
  geom_errorbar(aes(ymin= mh_prev_low, ymax= mh_prev_upp), 
                width=.2, position=position_dodge(.9))
```

One interesting observation is that same as depression, men who have children in the household have significantly lower prevalence of self-reported over 15 days of bad mental health while women who have children don't. We do a chi-square test incorporating the survey weights to verify this observation, below are the p-values for women and men:

```{r}
# For women
brfss_design %>%
  filter(sex == "female") %>% 
  select(children, mh_bin) %>% 
  svychisq(~children + mh_bin, . , statistic="adjWald") %>% 
  broom::tidy() %>% 
  pull(p.value)

# For men 
brfss_design %>%
  filter(sex == "male") %>% 
  select(children, mh_bin) %>% 
  svychisq(~children + mh_bin, . , statistic="adjWald") %>% 
  broom::tidy()%>% 
  pull(p.value)
```

As expected, based on the chi-squared test, the number of children in the household is significantly associated with self-reported over 15 days of bad mental health in the past 30 days among men but not women. 

### Gender, race, and # of days of bad mental health in the past 30 days

Next, we examine how gender and race together associate with self-reported prevalence of over 15 days of bad mental health in the past 30 days.

```{r}
brfss_design %>%
  group_by(sex, race, mh_bin) %>%
  summarize(mh_prev = 100*survey_mean(na.rm = T, vartype = c("ci"))) %>%
  filter(!is.na(race), mh_bin == ">15 days") %>%
  select(-mh_bin) %>%
  ggplot(aes(x = race, y = mh_prev, fill = sex)) +
  geom_bar(position="dodge", stat="identity") +
  geom_errorbar(aes(ymin=mh_prev_low, ymax=mh_prev_upp), 
                width=.2, position=position_dodge(.9))
```

We can see that not only is the prevalence of reported more than 15 days in the past 30 days being in bad mental health higher in women than men, the same holds true for all race/ethnicities included in the data set.

### Gender, age, and # of days of bad mental health in the past 30 days

Then, we want to examine the association between age and prevalence of more than 15 days being in bad mental health among different genders.

```{r}
brfss_design %>%
  group_by(sex, age, mh_bin) %>%
  summarize(mh_prev = 100*survey_mean(na.rm = T, vartype = c("ci"))) %>%
  filter(!is.na(age), mh_bin == ">15 days") %>%
  select(-mh_bin) %>%
  ggplot(aes(x = age, y = mh_prev, fill = sex)) +
  geom_bar(position="dodge", stat="identity") +
  geom_errorbar(aes(ymin=mh_prev_low, ymax=mh_prev_upp), 
                width=.2, position=position_dodge(.9))
```

Below are the p-values of the chi-square tests for the self-reported prevalence of more than 15 days of depression in the last 30 days in different age groups, for women and men separately.

```{r}
# For women
brfss_design %>%
  filter(sex == "female") %>% 
  select(age, mh_bin) %>% 
  svychisq(~age + mh_bin, . , statistic="adjWald") %>% 
  broom::tidy()%>% 
  pull(p.value)

# For men
brfss_design %>%
  filter(sex == "male") %>% 
  select(age, mh_bin) %>% 
  svychisq(~age + mh_bin, . , statistic="adjWald") %>% 
  broom::tidy()%>% 
  pull(p.value)
```

Based on the plot and test result above, it seems that, for women, the prevalence of self-reported over 15 days of bad mental in past 30 days drops slightly when entering the age group of 50-64, but decreases noticeably when entering 30-49 as well as 65+. The confidence interval plot below supported our conclusion. The same trend is displayed among men as well but in smaller magnitude.  

This gap in the magnitude of change in prevalence of self-reported number of days in bad mental in the past 30 days indicates that aging might affect self-reported it among women and men differently. 

```{r}
brfss_design %>%
  group_by(sex, age, mh_bin) %>%
  summarize(mh_prev = 100*survey_mean(na.rm = T, vartype = c("ci"))) %>%
  filter(!is.na(age), mh_bin == ">15 days") %>% 
  ggplot() + 
  geom_segment(aes(x = mh_prev_low, xend = mh_prev_upp,
                   y = age, yend = age, color = sex),
               size = 2)+
  labs(
    title = "Prevalence of over 15 days of Bad Mental Health during the Last 30 days",
    x = "Confidence Interval",
    y = "Age"
  ) 
```

### Conclusion

As a conclusion, all factors we chose influence self-reported number of days in bad mental in the past 30 days in the similar way to depression. We presume that it is because longer periods of being in bad mental health will lead to depression. Correspondingly, those in depression are much more likely to be in bad mental health for longer period.
