---
title: "Ecological Analysis"
author: ""
date: ""
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tidyverse)
library(ggplot2)
library(survey)
library(srvyr)

load(file = "./data/brfss_design.Rdata")

# Set options for allowing a single observation per stratum 
options(survey.lonely.psu = "adjust")
```

We obtained the U.S. Women, Peace and Security (WPS) Index in 2020 calculated by Georgetown Institute for Women, Peace, and Security downloaded from {here}(https://giwps.georgetown.edu/usa-index/). We import the raw data (xlsx) and extract the WPS index below:

```{r}
index_data_colnames = read_excel("./data/US-Index-Data.xlsx",
                        sheet = "Statistical table 3  Key statis",
                        n_max = 0) %>%
  names()

index_data = read_excel("./data/US-Index-Data.xlsx",
                        sheet = "Statistical table 3  Key statis",
                        range = "A3:L53",
                        col_names = index_data_colnames) %>% 
  janitor::clean_names()

index_data
```

We only need the overall WPS index and the states, so we select those two columns below:
```{r}
index_data = index_data %>% 
  select(state, index_score) %>%
  mutate(state_code = c(state.abb, "DC")[match(state, c(state.name, "District of Columbia"))],
         scaled_index = 100 * index_score)

index_data
```

We also obtained the per-capita GDP for each US state in 2021 from {here}(https://www.statista.com/statistics/248063/per-capita-us-real-gross-domestic-product-gdp-by-state/). We import the data below:
```{r}
gdp_data = read_excel("./data/us-per-capita-gdp-2021-by-state.xlsx",
                      sheet = "Data",
                      skip = 5,
                      col_names = c("state", "gdp")) %>% 
  mutate(state_code = c(state.abb, "DC")[match(state, c(state.name, "District of Columbia"))])

gdp_data
```

Next, we obtain the data frames on state-level self-reported depression prevalence data among women, and the state-level sex difference in self-repored depression prevalence, respectively:
```{r}
depr_prev_bystate_gender = brfss_design %>%
  group_by(state_code, sex, depression) %>%
  summarize(depression_prev = 100*survey_mean(na.rm = T, vartype = c("ci")))

depr_prev_bystate_female = depr_prev_bystate_gender %>%
  filter(sex == "female", depression == "yes") %>% 
  select(state_code, depression_prev)

depr_prev_bystate_female
```

```{r}
depr_prev_bystate_diff = depr_prev_bystate_gender %>%
  select(state_code, depression_prev, sex, depression) %>%
  filter(depression == "yes") %>%
  pivot_wider(names_from = "sex",
              values_from = "depression_prev") %>%
  mutate(difference = female - male) 

depr_prev_bystate_diff
```

Next, merge the WPS index data and the GDP data with the two above data frames, respectively:
```{r}
merged_female_depr_eco = list(depr_prev_bystate_female, index_data, gdp_data) %>% 
  reduce(left_join, by='state_code')

merged_female_depr_eco
```

The following scatterplot shows that there might be a negative association between the depression prevalence among women and the WPS index:
```{r}
merged_female_depr_eco %>% 
  ggplot(aes(x = scaled_index, y = depression_prev)) +
  geom_point()
```


We check if the association above is significant by running a multiple linear regression of Female depression prevalence in each state on the WPS index in each state:
```{r}
index_m11 = lm(depression_prev ~ scaled_index, data = merged_female_depr_eco)
summary(index_m11) %>% broom::tidy()
```
We check if the association between state-level prevalence of self-reported depression and the WPS index is still statistically significant after adjusting for state per-capita GDP:
```{r}
index_m12 = lm(depression_prev ~ scaled_index + gdp, data = merged_female_depr_eco)
summary(index_m12) %>% broom::tidy()
```

