---
title: "Regression modelling of self-reported depression diagnosis"
output: 
 html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(survey)
library(srvyr)
#library(svyVGAM)
library(readr)
library(forcats)
library(gtsummary)

# Set options for allowing a single observation per stratum 
options(survey.lonely.psu = "adjust")
```


```{r include = FALSE}
load(file = "./data/brfss_design.Rdata")
```

Based on results of former researches and EDA, we categorized variables selected into main exposures, demographic, socioeconomic, and health variables, and fitted model based on their category.

* Main exposures: sex, children
* Demographic variables: race, age and marital status 
* Socioeconomic variables: education, income, employment
* Health variables: general health, exercise

### Crude model with exposures only
 At the beginning, we constructed a crude model with these two exposures only:
*crude_mod_depr: depression ~ sex + children.* 
```{r}
crude_mod_depr = svyglm(depression ~ sex + children, # exposures
       brfss_design,
       family = quasibinomial())
crude_mod_depr %>% broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95) %>% knitr::kable(digits = 3)
```

### Model controlling for demographic variables
Then we wanted to take demographic covariates into account in our second model, including race, age and marital status: 
*mod2_depr: depression ~ sex + children + race + age + marital_status* 
```{r}
mod2_depr = svyglm(depression ~ sex + children + # exposures
         race + age + marital_status, # demographic covariates
       brfss_design,
       family = quasibinomial())
mod2_depr %>% broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95) %>% knitr::kable(digits = 3)
```

### Model controlling for demographic and socioeconomic variables
In the third model, we added socioeconomic covariates (education, income, employment) comparing to the second one:
*mod3_depr: depression ~ sex + children + race + age + marital_status + education + income + employment*
```{r}
mod3_depr = svyglm(depression ~ sex + children + # exposures
         race + age + marital_status + # demographic covariates
         education + income + employment, # socioeconomic covariates, 
       brfss_design,
       family = quasibinomial()) 
mod3_depr %>% broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95) %>% knitr::kable(digits = 3)
```


### Model controlling for demographic, socioeconomic, and health variables
Eventually, we fitted the full model considering all covariates:
*full_mod_depr: depression ~ sex + children + race + age + marital_status + education + income + employment + general_health + exercise*
```{r}
full_mod_depr = svyglm(depression ~ sex + children + # exposures
           race + age + marital_status + # demographic covariates
           education + income + employment + # socioeconomic covariates
           general_health + exercise, # health-related covariates
       brfss_design,
       family = quasibinomial())  

full_mod_depr %>% broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95) %>% knitr::kable(digits = 3)
```

### Model Comparison
We compared all these models based on Akaike Information Criterion (AIC), which is an estimator of prediction error and thereby relative quality of statistical models for a given set of data. Suppose that we have a statistical model of some data. Let p be the number of estimated parameters in the model. Let $\hat{L}$ be the maximized value of the likelihood function for the model. Then the AIC value of the model is the following: 

$AIC = 2k - 2ln(\hat{L})$.

Given a set of candidate models for the data, the preferred model is the one with the **minimum AIC value**. 
```{r, warning = FALSE}
getAIC = function(m){
  AIC(m) %>%
  broom::tidy() %>%
  filter(names == "AIC") %>%
    pull(x)}

variables = c("Exposures only", "Controlling for demographic variables", "Controlling for demographic and socioeconomic variables", "Controlling for demographic, socioeconomic, and health variables")
model_name = c("crude_mod_depr", "mod2_depr", "mod3_depr", "full_mod_depr")
AIC = c(getAIC(crude_mod_depr), getAIC(mod2_depr), getAIC(mod3_depr), getAIC(full_mod_depr))
df = data.frame(model_name, AIC, variables)

df %>%
  knitr::kable(digits = 3)
```

From the results of models comparison, we can conclude that the full model (model controlling for demographic, socioeconomic, and health variables) is the best model.

### Subsetted analysis by gender
Here we are interested in our second exposure, the number of children in the household. Specifically, we are interested in if it affects self-reported depression diagnosis differently for men and women. 

For male: 
```{r}
 brfss_male = brfss_design %>%
   filter(sex == "male")
 
 svyglm(depression ~ children + # exposures
            race + age + marital_status + # demographic covariates
            education + income + employment + # socioeconomic covariates
            general_health + exercise, # health-related covariates
        brfss_male,
        family = quasibinomial())  %>%
   broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95) %>%
  knitr::kable(digits = 3)
```

For female:
```{r}
 brfss_female = brfss_design %>%
   filter(sex == "female")
 
 svyglm(depression ~ children + # exposures
            race + age + marital_status + # demographic covariates
            education + income + employment + # socioeconomic covariates
            general_health + exercise, # health-related covariates
        brfss_female,
        family = quasibinomial())  %>%
   broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95)  %>%
   knitr::kable(digits = 3)
```

From subsetted regression, we can conclude that the number of children influences women and men in different ways because the p value for men now is much less than 0.05.






