---
title: "Regression modelling of number of days with bad mental health"
output: 
 html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(survey)
library(srvyr)
#library(svyVGAM)
library(readr)
library(forcats)
library(gtsummary)

# Set options for allowing a single observation per stratum 
options(survey.lonely.psu = "adjust")
```


```{r include = FALSE}
load(file = "./data/brfss_design.Rdata")
```

Based on results of former researches and EDA, we categorized variables selected into main exposures, demographic, socioeconomic, and health variables, and fitted model based on their category.

* Main exposures: sex, children
* Demographic variables: race, age and marital status 
* Socioeconomic variables: education, income, employment
* Health variables: general health, exercise

### Crude model with exposures only
We start with a crude model with the two exposures only.
```{r}
crude_mod_mh = svyglm(mh_bin ~ sex + children, # exposures
       brfss_design,
       family = quasibinomial()) 
crude_mod_mh %>%
  broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95) %>%
  knitr::kable(digits = 3)
```

### Model controlling for demographic variables
Then we wanted to take demographic covariates into account in our second model, including race, age and marital status. 
```{r}
mod2_mh = svyglm(mh_bin ~ sex + children + # exposures
         race + age + marital_status, # demographic covariates
       brfss_design,
       family = quasibinomial())
mod2_mh %>%
  broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95)  %>%
  knitr::kable(digits = 3)
```

### Model controlling for demographic and socioeconomic variables
In the third model, we added socioeconomic covariates (education, income, employment) to the second one.
```{r}
mod3_mh = svyglm(mh_bin ~ sex + children + # exposures
         race + age + marital_status + # demographic covariates
         education + income + employment, # socioeconomic covariates, 
       brfss_design,
       family = quasibinomial())
mod3_mh %>%
  broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95)  %>%
  knitr::kable(digits = 3)
```


### Full model 
Eventually, we fitted the full model controlling for all demographic, socioeconomic, and health variables:
```{r}
full_mod_mh = svyglm(mh_bin ~ sex + children + # exposures
           race + age + marital_status + # demographic covariates
           education + income + employment + # socioeconomic covariates
           general_health + exercise, # health-related covariates
       brfss_design,
       family = quasibinomial())  
full_mod_mh %>%
  broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95) %>%
  knitr::kable(digits = 3)
```

### Model Comparison
We compared all these models based on Akaike Information Criterion (AIC), which is an estimator of prediction error and thereby relative quality of statistical models for a given set of data. Suppose that we have a statistical model of some data. Let p be the number of estimated parameters in the model. Let $\hat{L}$ be the maximized value of the likelihood function for the model. Then the AIC value of the model is the following: 

$AIC = 2k - 2ln(\hat{L})$.

Given a set of candidate models for the data, the preferred model is the one with the **minimum AIC value**. 
```{r, warning = FALSE}
getAIC = function(m){
  AIC(m) %>%
  broom::tidy() %>%
  filter(names == "AIC") %>%
    pull(x)}

variables = c("Exposures only", "Controlling for demographic variables", "Controlling for demographic and socioeconomic variables", "Controlling for demographic, socioeconomic, and health variables")
model_name = c("crude_mod_mh", "mod2_mh", "mod3_mh", "full_mod_mh")
AIC = c(getAIC(crude_mod_mh), getAIC(mod2_mh), getAIC(mod3_mh), getAIC(full_mod_mh))
df = data.frame(model_name, AIC, variables)

df %>%
  knitr::kable(digits = 3)
```

From the results of models comparison, we can conclude that the full model (model controlling for demographic, socioeconomic, and health variables) is the best model. From the full model, we can see that, in the full sample, being female is associated with higher odds of reporting >15 days mental health not good in the past 30 days, and that having one or two children in the household is associated with lower odds of reporting >15 days mental health not good in the past 30 days. 

### Subsetted analysis by gender
Here we are interested in our second exposure, the number of children in the household. Specifically, we are interested in if it affects the self-reported number of days with bad mental health differently for men and women.  


The model fit for the male-only subset is as follows:
```{r}
 brfss_male = brfss_design %>%
   filter(sex == "male")
 
 svyglm(mh_bin ~ children + # exposures
            race + age + marital_status + # demographic covariates
            education + income + employment + # socioeconomic covariates
            general_health + exercise, # health-related covariates
        brfss_male,
        family = quasibinomial())  %>%
   broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95)  %>%
  knitr::kable(digits = 3)
```

The model fit for the female-only subset is as follows:
```{r}
 brfss_female = brfss_design %>%
   filter(sex == "female")
 
 svyglm(mh_bin ~ children + # exposures
            race + age + marital_status + # demographic covariates
            education + income + employment + # socioeconomic covariates
            general_health + exercise, # health-related covariates
        brfss_female,
        family = quasibinomial())  %>%
   broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95)  %>%
   knitr::kable(digits = 3)
```

From subsetted regression, we can conclude that the number of children in the household influences women and men in different ways. Specifically, among men, having children in the household is not significantly associated with the odds of self-reported >15 days mental health not good in the past 30 days. However, for women, having one or two children in the household (but not three or more) is significantly associated with lower odds of reporting >15 days mental health not good in the past 30 days. This might suggest that having one or two children in the household is beneficial to women's temperal mental health. However, causal relationships cannot be established.  

### Additional findings
According to the final model, we found that being non-white, older age, being married, higher annual income, being employed for salary, excellent physical health, and phisical exercise are associated with lower odds of self-reporting >15 days mental health not good in the past 30 days. 







