---
title: "Regression modelling of number of days with bad mental health"
output: 
 html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(survey)
library(srvyr)
#library(svyVGAM)
library(readr)
library(forcats)
library(gtsummary)

# Set options for allowing a single observation per stratum 
options(survey.lonely.psu = "adjust")
```


```{r include = FALSE}
load(file = "./data/brfss_design.Rdata")
```


### Crude model with exposures only
```{r}
crude_mod_mh = svyglm(mh_bin ~ sex + children, # exposures
       brfss_design,
       family = quasibinomial()) 
crude_mod_mh %>%
  broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95)
```

### Model controlling for demographic variables
```{r}
mod2_mh = svyglm(mh_bin ~ sex + children + # exposures
         race + age + marital_status, # demographic covariates
       brfss_design,
       family = quasibinomial())
mod2_mh %>%
  broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95)
```

### Model controlling for demographic and socioeconomic variables
```{r}
mod3_mh = svyglm(mh_bin ~ sex + children + # exposures
         race + age + marital_status + # demographic covariates
         education + income + employment, # socioeconomic covariates, 
       brfss_design,
       family = quasibinomial())
mod3_mh %>%
  broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95)
```


### Model controlling for demographic, socioeconomic, and health variables
```{r}
full_mod_mh = svyglm(mh_bin ~ sex + children + # exposures
           race + age + marital_status + # demographic covariates
           education + income + employment + # socioeconomic covariates
           general_health + exercise, # health-related covariates
       brfss_design,
       family = quasibinomial())  
full_mod_mh %>%
  broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95)
```

### Model Comparison
(AIC for score-based comparison for inference purposes and cross validation for prediction purpose?)
```{r, warning = FALSE}
getAIC = function(m){
  AIC(m) %>%
  broom::tidy() %>%
  filter(names == "AIC") %>%
    pull(x)}

variables = c("Exposures only", "Controlling for demographic variables", "Controlling for demographic and socioeconomic variables", "Controlling for demographic, socioeconomic, and health variables")
model_name = c("crude_mod_mh", "mod2_mh", "mod3_mh", "full_mod_mh")
AIC = c(getAIC(crude_mod_mh), getAIC(mod2_mh), getAIC(mod3_mh), getAIC(full_mod_mh))
df = data.frame(model_name, AIC, variables)

df %>%
  knitr::kable()
```

From the results of models comparison, we can conclude that the full model (model controlling for demographic, socioeconomic, and health variables) is the best model.

### Subsetted analysis by gender
Here we are intereted in our second exposure, the number of children in the household. Specifically, we are interested in if it affects the self-reported number of days with bad mental health days differently for men and women. 

For male: 
```{r}
 brfss_male = brfss_design %>%
   filter(sex == "male")
 
 svyglm(mh_bin ~ children + # exposures
            race + age + marital_status + # demographic covariates
            education + income + employment + # socioeconomic covariates
            general_health + exercise, # health-related covariates
        brfss_male,
        family = quasibinomial())  %>%
   broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95)
```

For female:
```{r}
 brfss_female = brfss_design %>%
   filter(sex == "female")
 
 svyglm(mh_bin ~ children + # exposures
            race + age + marital_status + # demographic covariates
            education + income + employment + # socioeconomic covariates
            general_health + exercise, # health-related covariates
        brfss_female,
        family = quasibinomial())  %>%
   broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95)
```

From subsetted regression, we can conclude that the number of children influences women and men in different ways because the p value for women now is much less than 0.05.







