---
title: "Regression modelling of self-reported depression diagnosis"
output: 
 html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(survey)
library(srvyr)
#library(svyVGAM)
library(readr)
library(forcats)
library(gtsummary)
```


```{r include = FALSE}
brfss21_data = read_csv("./data/cleaned_brfss21.csv") %>%
  mutate(mh_cat3 = factor(mh_cat3, levels = c("none", "<=15 days", ">15 days")),
         mh_bin = factor(mh_bin, levels = c("<=15 days", ">15 days")),
         depression = factor(depression, levels = c("no", "yes")),
         sex = factor(sex, levels = c("male", "female")),
         children = factor(children, levels = c("none", "one or two", "three or more")),
         state = factor(state),
         race = factor(race, levels = c("white", "black", "asian", "hispanic", "other")),
         marital_status = factor(marital_status, levels = c("married", "not married")),
         education = factor(education, levels = c("less than high school", "high school or some college", "bachelors or higher")),
         income = factor(income, levels = c("<=35000", "35000-75000", ">75000")),
         employment = factor(employment, levels = c("employed", "homemaker/student/retired", "unemployed")),
         age = factor(age, levels = c("18-29", "30-49", "50-64", "65+")),
         general_health = factor(general_health, levels = c("excellent", "very good/good", "fair/poor")),
         exercise = factor(exercise, levels = c("yes", "no")))
```

```{r include = FALSE}
# Set options for allowing a single observation per stratum 
options(survey.lonely.psu = "adjust")

# Create survey design 
brfss_design = as_survey_design(brfss21_data, id = 1, strata = ststr, weight = llcpwt)
```

### Crude model with exposures only
```{r}
svyglm(depression ~ sex + children, # exposures
       brfss_design,
       family = quasibinomial()) %>%
  broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95)
```

### Model controlling for demographic variables
```{r}
svyglm(depression ~ sex + children + # exposures
         race + age + marital_status, # demographic covariates
       brfss_design,
       family = quasibinomial()) %>%
  broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95)
```

### Model controlling for demographic and socioeconomic variables
```{r}
svyglm(depression ~ sex + children + # exposures
         race + age + marital_status + # demographic covariates
         education + income + employment, # socioeconomic covariates, 
       brfss_design,
       family = quasibinomial()) %>%
  broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95)
```


### Model controlling for demographic, socioeconomic, and health variables
```{r}
svyglm(depression ~ sex + children + # exposures
           race + age + marital_status + # demographic covariates
           education + income + employment + # socioeconomic covariates
           general_health + exercise, # health-related covariates
       brfss_design,
       family = quasibinomial())  %>%
  broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95)
```

### Subsetted analysis by gender
Here we are intereted in our second exposure, the number of children in the household. Specifically, we are interested in if it affects self-reported depression diagnosis differently for men and women. 









