---
title: "Regression modelling of self-reported depression diagnosis"
output: 
 html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(survey)
library(srvyr)
#library(svyVGAM)
library(readr)
library(forcats)
library(gtsummary)

# Set options for allowing a single observation per stratum 
options(survey.lonely.psu = "adjust")
```


```{r include = FALSE}
load(file = "./data/brfss_design.Rdata")
```

Based on results of former researches and EDA, we categorized variables selected into main exposures, demographic, socioeconomic, and health variables, and fitted model based on their category.

* Main exposures: sex, children
* Demographic variables: race, age and marital status 
* Socioeconomic variables: education, income, employment
* Health variables: general health, exercise

### Crude model with exposures only

First, we constructed a crude model with the two exposures only:

```{r}
crude_mod_depr = svyglm(depression ~ sex + children, # exposures
       brfss_design,
       family = quasibinomial())
crude_mod_depr %>% broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95) %>% knitr::kable(digits = 3)
```

### Model controlling for demographic variables
Then we wanted to take demographic covariates into account in our second model, including race, age and marital status: 
```{r}
mod2_depr = svyglm(depression ~ sex + children + # exposures
         race + age + marital_status, # demographic covariates
       brfss_design,
       family = quasibinomial())
mod2_depr %>% broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95) %>% knitr::kable(digits = 3)
```

### Model controlling for demographic and socioeconomic variables
In the third model, we added socioeconomic covariates (education, income, employment) comparing to the second one:
```{r}
mod3_depr = svyglm(depression ~ sex + children + # exposures
         race + age + marital_status + # demographic covariates
         education + income + employment, # socioeconomic covariates, 
       brfss_design,
       family = quasibinomial()) 
mod3_depr %>% broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95) %>% knitr::kable(digits = 3)
```


### Model controlling for demographic, socioeconomic, and health variables
Eventually, we fitted the full model considering all covariates:
```{r}
full_mod_depr = svyglm(depression ~ sex + children + # exposures
           race + age + marital_status + # demographic covariates
           education + income + employment + # socioeconomic covariates
           general_health + exercise, # health-related covariates
       brfss_design,
       family = quasibinomial())  

full_mod_depr %>% broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95) %>% knitr::kable(digits = 3)
```

### Model Comparison
We compared all these models based on Akaike Information Criterion (AIC), which is an estimator of prediction error and thereby relative quality of statistical models for a given set of data. Suppose that we have a statistical model of some data. Let p be the number of estimated parameters in the model. Let $\hat{L}$ be the maximized value of the likelihood function for the model. Then the AIC value of the model is the following: 

$AIC = 2k - 2ln(\hat{L})$.

Given a set of candidate models for the data, the preferred model is the one with the **minimum AIC value**. 
```{r, warning = FALSE}
getAIC = function(m){
  AIC(m) %>%
  broom::tidy() %>%
  filter(names == "AIC") %>%
    pull(x)}

variables = c("Exposures only", "Controlling for demographic variables", "Controlling for demographic and socioeconomic variables", "Controlling for demographic, socioeconomic, and health variables")
model_name = c("crude_mod_depr", "mod2_depr", "mod3_depr", "full_mod_depr")
AIC = c(getAIC(crude_mod_depr), getAIC(mod2_depr), getAIC(mod3_depr), getAIC(full_mod_depr))
df = data.frame(model_name, AIC, variables)

df %>%
  knitr::kable(digits = 3)
```

From the results of models comparison, we can conclude that the full model (model controlling for demographic, socioeconomic, and health variables) is the best model. We can see that being female is associated with higher odds of reporting past depression, adjusted for other covariates. 

### Subsetted analysis by gender
Here we are interested in our second exposure, the number of children in the household. Specifically, we are interested in if it affects self-reported depression diagnosis differently for men and women. In the full sample, the fitted model indicates that having children in the household is associated with lower odds of reporting depression. We subsetted the data into a men-only set and a women-only set and examine the effect of children separately. 

For men: 
```{r}
 brfss_male = brfss_design %>%
   filter(sex == "male")
 
 svyglm(depression ~ children + # exposures
            race + age + marital_status + # demographic covariates
            education + income + employment + # socioeconomic covariates
            general_health + exercise, # health-related covariates
        brfss_male,
        family = quasibinomial())  %>%
   broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95) %>%
  knitr::kable(digits = 3)
```

For women:
```{r}
 brfss_female = brfss_design %>%
   filter(sex == "female")
 
 svyglm(depression ~ children + # exposures
            race + age + marital_status + # demographic covariates
            education + income + employment + # socioeconomic covariates
            general_health + exercise, # health-related covariates
        brfss_female,
        family = quasibinomial())  %>%
   broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95)  %>%
   knitr::kable(digits = 3)
```

From subsetted regression, we can conclude that the number of children influences women and men in different ways. Specifically, for men, having children in the household is associated with lower odds of reporting depression. However, for women, having children in the household is not associated with lower odds of reporting depression. 

### Additional findings
According to the final model, we found that being non-white, older age, being married, higher annual income, being employed for salary, excellent physical health, and phisical exercise are associated with lower odds of self-reported depression. 







