---
title: "Regression modelling of self-reported depression diagnosis"
output: 
 html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(survey)
library(srvyr)
#library(svyVGAM)
library(readr)
library(forcats)
library(gtsummary)

# Set options for allowing a single observation per stratum 
options(survey.lonely.psu = "adjust")
```


```{r include = FALSE}
load(file = "./data/brfss_design.Rdata")
```


### Crude model with exposures only
```{r}
crude_mod_depr = svyglm(depression ~ sex + children, # exposures
       brfss_design,
       family = quasibinomial())
crude_mod_depr %>% broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95)
```

### Model controlling for demographic variables
```{r}
mod2_depr = svyglm(depression ~ sex + children + # exposures
         race + age + marital_status, # demographic covariates
       brfss_design,
       family = quasibinomial())
mod2_depr %>% broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95)
```

### Model controlling for demographic and socioeconomic variables
```{r}
mod3_depr = svyglm(depression ~ sex + children + # exposures
         race + age + marital_status + # demographic covariates
         education + income + employment, # socioeconomic covariates, 
       brfss_design,
       family = quasibinomial()) 
mod3_depr %>% broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95)
```


### Model controlling for demographic, socioeconomic, and health variables
```{r}
full_mod_depr = svyglm(depression ~ sex + children + # exposures
           race + age + marital_status + # demographic covariates
           education + income + employment + # socioeconomic covariates
           general_health + exercise, # health-related covariates
       brfss_design,
       family = quasibinomial())  

full_mod_depr %>% broom::tidy(exponentiate=T, conf.int = T, conf.level=0.95)
```

### Model Comparison
(AIC for score-based comparison for inference purposes and cross validation for prediction purpose?)
```{r}
AIC(full_mod_depr)
```


### Subsetted analysis by gender
Here we are intereted in our second exposure, the number of children in the household. Specifically, we are interested in if it affects self-reported depression diagnosis differently for men and women. 









