---
title: "State-level mental health outcome by gender"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    source: embed
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(ggplot2)
library(survey)
library(srvyr)
library(readr)
library(forcats)
library(plotly)
```

```{r include = FALSE}
brfss21_data = read_csv("./data/cleaned_brfss21.csv") %>%
  mutate(mh_cat3 = factor(mh_cat3, levels = c("none", "<=15 days", ">15 days")),
         mh_bin = factor(mh_bin, levels = c("<=15 days", ">15 days")),
         depression = factor(depression, levels = c("no", "yes")),
         sex = factor(sex, levels = c("male", "female")),
         children = factor(children, levels = c("none", "one or two", "three or more")),
         state = factor(state),
         race = factor(race, levels = c("white", "black", "asian", "hispanic", "other")),
         education = factor(education, levels = c("less than high school", "high school or some college", "bachelors or higher")),
         income = factor(income, levels = c("<=35000", "35000-75000", ">75000")),
         employment = factor(employment, levels = c("employed", "homemaker/student/retired", "unemployed")),
         age = factor(age, levels = c("18-29", "30-49", "50-64", "65+")),
         general_health = factor(general_health, levels = c("excellent", "very good/good", "fair/poor")),
         exercise = factor(exercise, levels = c("yes", "no")))
```

```{r include = FALSE}
# Set options for allowing a single observation per stratum 
options(survey.lonely.psu = "adjust")

# Create survey design 
brfss_design = as_survey_design(brfss21_data, id = 1, strata = ststr, weight = llcpwt)
```

```{r include = FALSE}
depr_prev_bystate_gender = brfss_design %>%
  group_by(state_code, sex, depression) %>%
  summarize(depression_prev = 100*survey_mean(na.rm = T, vartype = c("ci")))
  #summarize(depression_prev = 100*survey_mean(na.rm = T, vartype = c("ci"), proportion = T))

depr_prev_bystate_f = depr_prev_bystate_gender %>%
  filter(sex == "female", depression == "yes")

depr_prev_bystate_m = depr_prev_bystate_gender %>%
  filter(sex == "male", depression == "yes")

# construct new data frame for plotting
depr_prev_bystate_diff = depr_prev_bystate_gender %>%
  select(state_code, depression_prev, sex, depression) %>%
  filter(depression == "yes") %>%
  pivot_wider(names_from = "sex",
              values_from = "depression_prev") %>%
  mutate(difference = female - male)

```

```{r include = FALSE}
days_bad_mh_bystate_gender = brfss_design %>%
  group_by(state_code, sex) %>%
  summarize(days = survey_mean(days_bad_mental_health, na.rm = T, vartype = c("ci")))

days_bad_mh_bystate_f = days_bad_mh_bystate_gender %>%
  filter(sex == "female")

days_bad_mh_bystate_m = days_bad_mh_bystate_gender %>%
  filter(sex == "male")

# construct new data frame for plotting
days_bad_mh_bystate_diff = days_bad_mh_bystate_gender %>%
  select(state_code, sex, days) %>%
  pivot_wider(names_from = "sex",
              values_from = "days") %>%
  mutate(difference = female - male)
```


### State-level prevalence of self-reported past depression by gender

#### Prevalence of self-reported past depression (%) among women

```{r}
depr_prev_bystate_f_graph = plot_geo(depr_prev_bystate_f, 
                                     locationmode = "USA-states",
                                     text = ~paste('95% CI: (', 
                                                   round(depression_prev_low,3),
                                                   round(depression_prev_upp,3),
                                                   ")")) %>%
  add_trace(locations = ~state_code,
            z = ~depression_prev,
            color = ~depression_prev,
            colors = "Purples") %>%
  layout(geo = list(scope = "usa")) %>%
         #legend = list(title = list(text = "<b>Cylinders</b>")),
         #title = "Prevalence of diagnosis of depression (%) among women") %>%
  colorbar(ticksuffix = "%") 


depr_prev_bystate_f_graph
```


#### Prevalence of self-reported past depression (%) among men

```{r}
depr_prev_bystate_m_graph = plot_geo(depr_prev_bystate_m, 
                                     locationmode = "USA-states",
                                     text = ~paste('95% CI: (', 
                                                   round(depression_prev_low,3),
                                                   round(depression_prev_upp,3),
                                                   ")")) %>%
  add_trace(locations = ~state_code,
            z = ~depression_prev,
            color = ~depression_prev,
            colors = "Purples") %>%
  layout(geo = list(scope = "usa")) %>%
         #legend = list(title = list(text = "<b>Cylinders</b>")),
  colorbar(ticksuffix = "%") 


depr_prev_bystate_m_graph
```

#### Gender difference in prevalence (%) of self-reported past depression (F minus M)

```{r}
depr_prev_bystate_diff_graph = plot_geo(depr_prev_bystate_diff, 
                                     locationmode = "USA-states",
                                     text = ~paste("Difference in prevalence: ",
                                                   round(difference,3))) %>%
  add_trace(locations = ~state_code,
            z = ~round(difference,3),
            color = ~difference,
            colors = "Purples") %>%
  layout(geo = list(scope = "usa")) %>%
  colorbar(ticksuffix = "%") 


depr_prev_bystate_diff_graph
```

***

In summary, states in the South (e.g., West Virginia, Kentucky, Tennessee), Northwest (e.g, Washington, Oregon), and North (e.g., Maine, New Hampshire) have the highest prevalence of self-reported past depression among women. The same pattern in geographic distribution holds in men, but they have lower prevalence of self-reported past depression than women. The gender difference in prevalence of self-reported past depression is the largest in the South, West (except California, Nevada, and Arizona) and the Midwest (except Illinois). 


### State-level average self-reported number of days mental health not good in the past 30 days by gender

#### Average number of days of mental health not good in the past 30 days among women

```{r}
avg_days_bad_mh_bystate_f_graph = plot_geo(days_bad_mh_bystate_f, 
                                     locationmode = "USA-states",
                                     text = ~paste('95% CI: (', 
                                                   round(days_low, 3),
                                                   round(days_upp, 3),
                                                   ")")) %>%
  add_trace(locations = ~state_code,
            z = ~round(days, 3),
            color = ~days,
            colors = "Purples") %>%
  layout(geo = list(scope = "usa")) 

avg_days_bad_mh_bystate_f_graph
```

#### Average number of days mental health not good in the past 30 days among men
```{r}
avg_days_bad_mh_bystate_m_graph = plot_geo(days_bad_mh_bystate_m, 
                                     locationmode = "USA-states",
                                     text = ~paste('95% CI: (', 
                                                   round(days_low, 3),
                                                   round(days_upp, 3),
                                                   ")")) %>%
  add_trace(locations = ~state_code,
            z = ~round(days, 3),
            color = ~days,
            colors = "Purples") %>%
  layout(geo = list(scope = "usa")) 

avg_days_bad_mh_bystate_m_graph
```

#### Gender difference in average number of days mental health not good in the past 30 days (F minus M)
```{r}
avg_day_bad_mh_bystate_diff_graph = plot_geo(days_bad_mh_bystate_diff, 
                                     locationmode = "USA-states",
                                     text = ~paste("Difference in average days: ",
                                                   round(difference,3))) %>%
  add_trace(locations = ~state_code,
            z = ~round(difference,3),
            color = ~difference,
            colors = "Purples") %>%
  layout(geo = list(scope = "usa")) 


avg_day_bad_mh_bystate_diff_graph
```

***

In summary, women in Southern states report the highest average number of days mental health not good in the past 30 days. Similar patterns hold for men, but they on average report fewer bad mental health days. The gender difference is the largest in the South and the Midwest. 