---
title: "State-level mental health outcome by gender"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    source: embed
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(ggplot2)
library(survey)
library(srvyr)
library(readr)
library(forcats)
library(plotly)
```

```{r include = FALSE}
brfss21_data = read_csv("./data/cleaned_brfss21.csv") %>%
  mutate(mh_cat3 = factor(mh_cat3, levels = c("none", "<=15 days", ">15 days")),
         mh_bin = factor(mh_bin, levels = c("<=15 days", ">15 days")),
         depression = factor(depression, levels = c("no", "yes")),
         sex = factor(sex, levels = c("male", "female")),
         children = factor(children, levels = c("none", "one or two", "three or more")),
         state = factor(state),
         race = factor(race, levels = c("white", "black", "asian", "hispanic", "other")),
         education = factor(education, levels = c("less than high school", "high school or some college", "bachelors or higher")),
         income = factor(income, levels = c("<=35000", "35000-75000", ">75000")),
         employment = factor(employment, levels = c("employed", "homemaker/student/retired", "unemployed")),
         age = factor(age, levels = c("18-29", "30-49", "50-64", "65+")),
         general_health = factor(general_health, levels = c("excellent", "very good/good", "fair/poor")),
         exercise = factor(exercise, levels = c("yes", "no")))
```

```{r include = FALSE}
# Set options for allowing a single observation per stratum 
options(survey.lonely.psu = "adjust")

# Create survey design 
brfss_design = as_survey_design(brfss21_data, id = 1, strata = ststr, weight = llcpwt)
```

```{r include = FALSE}
depr_prev_bystate_gender = brfss_design %>%
  group_by(state_code, sex, depression) %>%
  summarize(depression_prev = 100*survey_mean(na.rm = T, vartype = c("ci")))
  #summarize(depression_prev = 100*survey_mean(na.rm = T, vartype = c("ci"), proportion = T))

depr_prev_bystate_f = depr_prev_bystate_gender %>%
  filter(sex == "female", depression == "yes")

depr_prev_bystate_m = depr_prev_bystate_gender %>%
  filter(sex == "male", depression == "yes")

# construct new data frame for plotting
depr_prev_bystate_diff = depr_prev_bystate_gender %>%
  select(state_code, depression_prev, sex, depression) %>%
  filter(depression == "yes") %>%
  pivot_wider(names_from = "sex",
              values_from = "depression_prev") %>%
  mutate(difference = female - male)

```

```{r include = FALSE}
days_bad_mh_bystate_gender = brfss_design %>%
  group_by(state_code, sex) %>%
  summarize(days = survey_mean(days_bad_mental_health, na.rm = T, vartype = c("ci")))

days_bad_mh_bystate_f = days_bad_mh_bystate_gender %>%
  filter(sex == "female")

days_bad_mh_bystate_m = days_bad_mh_bystate_gender %>%
  filter(sex == "male")

# construct new data frame for plotting
days_bad_mh_bystate_diff = days_bad_mh_bystate_gender %>%
  select(state_code, sex, days) %>%
  pivot_wider(names_from = "sex",
              values_from = "days") %>%
  mutate(difference = female - male)
```


### State-level prevalence of self-reported past diagnosis of depression by gender

#### Prevalence of self-reported past diagnosis of depression (%) among women

```{r}
depr_prev_bystate_f_graph = plot_geo(depr_prev_bystate_f, 
                                     locationmode = "USA-states",
                                     text = ~paste('95% CI: (', 
                                                   round(depression_prev_low,3),
                                                   round(depression_prev_upp,3),
                                                   ")")) %>%
  add_trace(locations = ~state_code,
            z = ~depression_prev,
            color = ~depression_prev,
            colors = "Purples") %>%
  layout(geo = list(scope = "usa")) %>%
         #legend = list(title = list(text = "<b>Cylinders</b>")),
         #title = "Prevalence of diagnosis of depression (%) among women") %>%
  colorbar(ticksuffix = "%") 


depr_prev_bystate_f_graph
```


#### Prevalence of self-reported past diagnosis of depression (%) among men

```{r}
depr_prev_bystate_m_graph = plot_geo(depr_prev_bystate_m, 
                                     locationmode = "USA-states",
                                     text = ~paste('95% CI: (', 
                                                   round(depression_prev_low,3),
                                                   round(depression_prev_upp,3),
                                                   ")")) %>%
  add_trace(locations = ~state_code,
            z = ~depression_prev,
            color = ~depression_prev,
            colors = "Purples") %>%
  layout(geo = list(scope = "usa")) %>%
         #legend = list(title = list(text = "<b>Cylinders</b>")),
  colorbar(ticksuffix = "%") 


depr_prev_bystate_m_graph
```

#### Gender difference in prevalence (%) of self-reported past diagnosis of depression (F minus M)

```{r}
depr_prev_bystate_diff_graph = plot_geo(depr_prev_bystate_diff, 
                                     locationmode = "USA-states",
                                     text = ~paste("Difference in prevalence: ",
                                                   round(difference,3))) %>%
  add_trace(locations = ~state_code,
            z = ~round(difference,3),
            color = ~difference,
            colors = "Purples") %>%
  layout(geo = list(scope = "usa")) %>%
  colorbar(ticksuffix = "%") 


depr_prev_bystate_diff_graph
```

***

INSERT COMMENTS/TAKEAWAYS HERE 


### State-level average self-reported number of days mental health not good in the past 30 days by gender

#### Average number of days of self-reported bad mental health in the past 30 days among women

```{r}
avg_days_bad_mh_bystate_f_graph = plot_geo(days_bad_mh_bystate_f, 
                                     locationmode = "USA-states",
                                     text = ~paste('95% CI: (', 
                                                   round(days_low, 3),
                                                   round(days_upp, 3),
                                                   ")")) %>%
  add_trace(locations = ~state_code,
            z = ~round(days, 3),
            color = ~days,
            colors = "Purples") %>%
  layout(geo = list(scope = "usa")) 

avg_days_bad_mh_bystate_f_graph
```

#### Average number of days of self-reported bad mental health in the past 30 days among men
```{r}
avg_days_bad_mh_bystate_m_graph = plot_geo(days_bad_mh_bystate_m, 
                                     locationmode = "USA-states",
                                     text = ~paste('95% CI: (', 
                                                   round(days_low, 3),
                                                   round(days_upp, 3),
                                                   ")")) %>%
  add_trace(locations = ~state_code,
            z = ~round(days, 3),
            color = ~days,
            colors = "Purples") %>%
  layout(geo = list(scope = "usa")) 

avg_days_bad_mh_bystate_m_graph
```

#### Gender difference in average number of days of self-reported bad mental health in the past 30 days (F minus M)
```{r}
avg_day_bad_mh_bystate_diff_graph = plot_geo(days_bad_mh_bystate_diff, 
                                     locationmode = "USA-states",
                                     text = ~paste("Difference in average days: ",
                                                   round(difference,3))) %>%
  add_trace(locations = ~state_code,
            z = ~round(difference,3),
            color = ~difference,
            colors = "Purples") %>%
  layout(geo = list(scope = "usa")) 


avg_day_bad_mh_bystate_diff_graph
```

***

INSERT COMMENTS/TAKEAWAYS HERE 

### Side-by-side comparisons

#### Prevalence of self-reported past diagnosis of depression by gender in each US state

```{r}
grouped_depr_prev_barplot = depr_prev_bystate_gender %>%
  filter(depression == "yes") %>%
  plot_ly(y = ~state_code, x = ~depression_prev, 
          color = ~sex, type = "bar", colors = "viridis") %>%
  layout(xaxis = list(title = 'prevalence (%)'), 
         yaxis = list(title = 'state'))

grouped_depr_prev_barplot

```

#### Average number of days of self-reported bad mental health in the past 30 days by gender in each US state

```{r}
grouped_avg_days_bad_mh_barplot = days_bad_mh_bystate_gender %>%
  plot_ly(y = ~state_code, x = ~days, 
          color = ~sex, type = "bar", colors = "viridis") %>%
  layout(xaxis = list(title = '# of days'), 
         yaxis = list(title = 'state'))

grouped_avg_days_bad_mh_barplot

```

***

INSERT COMMENTS/TAKEAWAYS HERE 
